Kids LRTI tracking

### 2020-08-25 lrti_tracking.Rmd
lrti data dates

@@

demographics <- read_rds(paste0(file_path, "Output/demographics_final.rds")) %>% 
  select(child_id_num, child_id_char, area, winter_id)
```

@@

  filter(!is.na(home)) %>% 
  filter(!is.na(paul_lrti)) %>%

@@

  select(area, home, child_id_num, home_winter_id, winter_id, 
         paul_lrti, paul_comments,
         visit_date:treatment_assigned) %>% 
  mutate(child_id_num = as.numeric(child_id_num),
         home_winter_id = as.numeric(home_winter_id)) %>% 
  arrange(area, child_id_num, home_winter_id)

@@

The filtering below narrows down the LRTI definition too much; we are missing cases
where children had LRTI but did not visit a healthcare facility and have a med rec.
If we only use these cases, it could bias the results.

There are instances of LRTI in our records that weren’t confirmed in the medical record. 
There are other instances where the medical record diagnosed LRTI, but our records 
overturned it. What is our criteria for diagnosing LRTI, where is it documented, 
and how can we be sure it is consistent in diagnosing LRTI?

```{r}
# Further filtering and re-ordering of variables
lrti_diagnosis_data_filtered <- lrti_diagnosis_data %>% 
  select(area:med_visit_ari, csq_date:csq_symptom_free_date, 
         med_rec_visit_date:med_rec_comments, med_visit_lethargy:csq_comments,
         lrti_number:lrti_professional_comments) %>% 
  group_by(area, home, home_winter_id, winter_id, child_id_num, visit_date) %>% 
  arrange(area, home, home_winter_id, winter_id, child_id_num, visit_date, csq_collect) %>% 
  # If a home CSQ was done on same visit date as initial phone CSQ, filter for only the home CSQ
  distinct(csq_sick_2_weeks, .keep_all = TRUE) %>% 
  select(area:csq_still_sick, med_visit_referred, csq_see_dr, 
         csq_first_symptom_date:lrti_professional_comments) %>% 
  # Were they referred (med visit) or did they see a Dr (CSQ)?
  filter(med_visit_referred == "Yes" | csq_see_dr == "Yes") %>% 
  # If yes, did we get medical records?
  filter(!is.na(med_rec_visit_date)) %>% 
  # If yes, was there an ARI diagnosis?
  filter(med_rec_ari == "ARI" | med_rec_ari == "Possible ARI") %>% 
  group_by(child_id_num) %>% 
  distinct(med_rec_visit_date, .keep_all = TRUE)
```

# Defining start date, end date, and at-risk period for each diagnosed LRTI case

First, I need to know (or make) exactly which variable is a final diagnosed case 
of LRTI in the data. Then, I can create a method in R to identify start/end/at-risk.
The code below is a start at this, using Nick's handoff documents/algorithm.

The initial attempt at having R select start dates did not match up well with
the cases in which Paul had already finalized the start/end dates. There seems
to be a good amount of professional opinion by assessing symptoms and medical
records. It may be best to filter a dataset down to only confirmed LRTI cases
and useful health/medical variables, then decide start/end/at-risk times 
on a case-by-case basis.

```{r}
# LRTI start date, end date, and at-risk period identification
lrti_date_data <- health_exposure_data_sampling_day %>% 
  select(area, home, home_winter_id, winter_id, child_id_num, visit_date,
         med_visit_date:med_visit_notes, chw_dr_relook:chw_ari_comments,
         csq_date:csq_comments, med_rec_visit_date:med_rec_comments,
         lrti_number:lrti_professional_comments) %>% 
  filter(area == "WMT") %>% 
  filter(csq_sick_2_weeks == "Yes") %>% 
  select(area:med_visit_ari, csq_date:csq_symptom_free_date, 
         med_rec_visit_date:med_rec_comments, 
         med_visit_lethargy:lrti_professional_comments) %>% 
  group_by(area, home, home_winter_id, winter_id, child_id_num, visit_date) %>% 
  arrange(area, home, home_winter_id, winter_id, child_id_num, visit_date, csq_collect) %>% 
  distinct(csq_sick_2_weeks, .keep_all = TRUE) %>% 
  select(area:csq_still_sick, med_visit_referred, csq_see_dr, 
         csq_first_symptom_date:lrti_professional_comments) %>% 
  mutate(trash_date = ymd("2099-01-01"),
         still_sick_date = if_else(csq_still_sick == "Yes", 
                                   csq_date, trash_date),
         med_rec_ari_date = if_else(!is.na(med_rec_visit_date), 
                                    med_rec_visit_date, trash_date),
         med_visit_ari_date = if_else(!is.na(med_visit_date), 
                                      med_visit_date, trash_date)) %>% 
  pivot_longer(cols = c("still_sick_date", "med_rec_ari_date", "med_visit_ari_date"),
               names_to = "lrti_visits", values_to = "lrti_dates") %>% 
  group_by(area, home, home_winter_id, winter_id, child_id_num, visit_date) %>% 
  mutate(lrti_start_date_new = if_else(!is.na(csq_first_symptom_date),
                                       csq_first_symptom_date, lrti_dates)) %>% 
  select(area, home, home_winter_id, winter_id, child_id_num, visit_date,
         csq_first_symptom_date, lrti_start_date_new, lrti_visits, lrti_dates, trash_date,
         csq_sick_2_weeks, csq_still_sick, med_rec_ari, med_visit_ari,
         lrti_number:lrti_professional_comments) %>% 
  arrange(area, home, home_winter_id, winter_id, child_id_num, visit_date, lrti_dates) %>% 
  distinct(visit_date, .keep_all = TRUE) %>% 
  group_by(area, home, home_winter_id, winter_id, child_id_num) %>% 
  distinct(lrti_start_date_new, .keep_all = TRUE) %>% 
  distinct(child_id_num, lrti_number, .keep_all = TRUE) %>% 
  filter(!is.na(lrti_number)) %>% 
  select(child_id_num, home_winter_id, lrti_number, lrti_start_date_new:lrti_dates, 
         lrti_start_date_final, lrti_end_date_final, area:lrti_professional_comments)
```

>>>>

check_data <- health_exposure_data_sampling_day %>% 
  filter(child_id_num == 68 & area == "NN")
  filter(home == "CH110")

@@

demographics <- read_rds(paste0(file_path, "Output/demographics_final.rds")) %>% 
  filter(child_id_num == 52)
  select(child_id_num, child_id_char, area, winter_id)
```

@@

  filter(!is.na(home)) %>% 
  select(area, home, child_id_num, home_winter_id, winter_id, 
         paul_lrti, paul_comments,
         visit_date:treatment_assigned) %>% 
  mutate(lrti_any_winter = if_else(!is.na(paul_lrti), 1, 0)) %>% 
  group_by(home) %>% 
  mutate(lrti_any_winter = sum(lrti_any_winter)) %>% 
  select(area, home, lrti_any_winter, child_id_num, home_winter_id, winter_id, 
         paul_lrti, paul_comments,
         visit_date:treatment_assigned) %>% 
  filter(lrti_any_winter > 0) %>%

@@

  mutate(child_id_num = as.numeric(child_id_num),
         home_winter_id = as.numeric(home_winter_id)) %>% 
  arrange(area, child_id_num, home_winter_id)

@@

# Merge dataset with final LRTI notes and dates, then assign person-time
```{r}
file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")

lrti_notes_decisions <- read_xlsx(paste0(file_path, "Input/lrti_notes_decisions.xlsx")) %>% 
  filter(lrti_final == "Yes") %>% 
  select(area, home, child_id_num, home_winter_id, winter_id, lrti_final, 
         lrti_start_date_final, lrti_end_date_final, notes_final, total_cases_this_winter)

wmt_homewinter_initial <- read_csv(paste0(file_path, "Input/WMT/homewinter.csv"), na = "NULL") %>% 
  rename_all(tolower) %>% 
  select(homewinterid, homeid, winterid, interventiondate, interventiontime) %>% 
  filter(winterid == 1 | winterid ==2) %>% # filter for winterid == 1 or 2. What about missing intervention date?
  mutate(home_winter_id = as.character(homewinterid),
         intervention_date = ymd(interventiondate),
         area = "WMT") %>% 
  select(area, home_winter_id, intervention_date)

nn_homewinter_initial <- read_csv(paste0(file_path, "Input/NN/homewinter.csv"), na = "NULL") %>% 
  rename_all(tolower) %>% 
  select(homewinterid, homeid, winterid, interventiondate, interventiontime) %>% 
  filter(winterid == 1 | winterid ==2) %>% # filter for winterid == 1 or 2. What about missing intervention date?
  mutate(home_winter_id = as.character(homewinterid),
         intervention_date = mdy(interventiondate),
         area = "NN") %>% 
  select(area, home_winter_id, intervention_date)

ak_homewinter_initial <- read_csv(paste0(file_path, "Input/AK/homewinter.csv"), na = "NULL") %>% 
  rename_all(tolower) %>% 
  select(homewinterid, homeid, winterid, interventiondate, interventiontime) %>% 
  filter(winterid == 1 | winterid ==2) %>% # filter for winterid == 1 or 2. What about missing intervention date?
  mutate(home_winter_id = as.character(homewinterid),
         intervention_date = mdy(interventiondate),
         area = "AK") %>% 
  select(area, home_winter_id, intervention_date)

intervention_dates <- rbind(wmt_homewinter_initial, 
                            nn_homewinter_initial,
                            ak_homewinter_initial)

lrti_yes <- lrti_notes_decisions %>% 
  mutate(child_id_num = as.character(child_id_num),
         home_winter_id = as.character(home_winter_id),
         winter_id = as.character(winter_id)) %>% 
  # join lrti data with full dataset
  full_join(health_exposure_data_sampling_day, 
            by = c("area", "home", "child_id_num", "home_winter_id", "winter_id")) %>% 
  filter(lrti_final == "Yes") %>% 
  # select vars for calculating person-time
  select(area, home, child_id_num, home_winter_id, winter_id, lrti_final, 
         lrti_start_date_final, lrti_end_date_final, notes_final, total_cases_this_winter,
         csq_date) %>% 
  left_join(intervention_dates, by = c("area", "home_winter_id")) %>% 
  group_by(area, child_id_num, winter_id) %>% 
  arrange(area, child_id_num, winter_id, csq_date) %>% 
  # var for person-time start date for each winter/child combo
  # this date is intervention date + 14 days
  mutate(person_time_start = intervention_date + 14) %>% 
  filter(!is.na(csq_date)) %>% 
  # var for person-time end date for each winter/child combo
  # this date is the earliest of the last csq or april 30 of the study year
  mutate(person_time_end = last(csq_date),
         latest_date = "04-30") %>% 
  separate(person_time_end, into = c("pte_year", "pte_mo", "pte_day"), sep = "-", remove = FALSE) %>% 
  unite(latest_date, c("pte_year", "latest_date"), sep = "-") %>% 
  select(-pte_mo, -pte_day) %>% 
  mutate(latest_date = ymd(latest_date),
         person_time_end = if_else(person_time_end > latest_date,
                                   latest_date, person_time_end),
         person_time_possible = person_time_end - person_time_start) %>% 
  # person-time possible is number of days from start to end date for each child/winter combo
  group_by(area, child_id_num, winter_id, lrti_start_date_final) %>% 
  # end of lrti is the lrti end date +14 days
  mutate(lrti_end_date_plus14 = ymd(lrti_end_date_final) + 14,
         lrti_end_date_plus14 = if_else(lrti_end_date_plus14 > latest_date,
                                 latest_date, lrti_end_date_plus14)) %>% 
  # filter out lrti cases that start later than person-time end date (2 cases)
  filter(lrti_start_date_final <= person_time_end) %>% 
  # calculate days with LRTI
  mutate(person_time_lrti = ymd(lrti_end_date_plus14) - ymd(lrti_start_date_final)) %>% 
  distinct(area, child_id_num, winter_id, lrti_start_date_final, .keep_all = TRUE) %>% 
  group_by(area, child_id_num, winter_id) %>% 
  # calculate total LRTI days per child/winter combo
  mutate(person_time_lrti_total = sum(person_time_lrti)) %>% 
  ungroup() %>% 
  #distinct(area, child_id_num, winter_id, .keep_all = TRUE) %>% 
  # new var for time without lrti
  mutate(person_time_no_lrti_total = person_time_possible - person_time_lrti_total) %>% 
  arrange(area, child_id_num, winter_id)

lrti_no <- lrti_notes_decisions %>% 
  mutate(child_id_num = as.character(child_id_num),
         home_winter_id = as.character(home_winter_id),
         winter_id = as.character(winter_id)) %>% 
  # join lrti data with full dataset
  full_join(health_exposure_data_sampling_day, 
            by = c("area", "home", "child_id_num", "home_winter_id", "winter_id")) %>% 
  filter(is.na(lrti_final)) %>% 
  # select vars for calculating person-time
  select(area, home, child_id_num, home_winter_id, winter_id, lrti_final, 
         lrti_start_date_final, lrti_end_date_final, notes_final, total_cases_this_winter,
         csq_date) %>% 
  left_join(intervention_dates, by = c("area", "home_winter_id")) %>% 
  group_by(area, child_id_num, winter_id) %>% 
  arrange(area, child_id_num, winter_id, csq_date) %>% 
  # var for person-time start date for each winter/child combo
  # this date is intervention date + 14 days
  mutate(person_time_start = intervention_date + 14) %>% 
  filter(!is.na(csq_date)) %>% 
  # var for person-time end date for each winter/child combo
  # this date is the earliest of the last csq or april 30 of the study year
  mutate(person_time_end = last(csq_date),
         latest_date = "04-30") %>% 
  separate(person_time_end, into = c("pte_year", "pte_mo", "pte_day"), sep = "-", remove = FALSE) %>% 
  unite(latest_date, c("pte_year", "latest_date"), sep = "-") %>% 
  select(-pte_mo, -pte_day) %>% 
  mutate(latest_date = ymd(latest_date),
         person_time_end = if_else(person_time_end > latest_date,
                                   latest_date, person_time_end),
         person_time_end = if_else(person_time_end < person_time_start, last(csq_date), person_time_end),
         person_time_possible = as.numeric(person_time_end - person_time_start),
         person_time_possible = if_else(person_time_possible < 0, 0,
                                        person_time_possible)) %>% 
  distinct(area, child_id_num, winter_id, .keep_all = TRUE) %>% 
  arrange(area, child_id_num, winter_id)

lrti_final_data <- plyr::rbind.fill(lrti_yes, lrti_no) %>% 
  arrange(area, child_id_num, winter_id)

write_csv(lrti_final_data, paste0(file_path, "Output/lrti_final_data.csv"), na = "")
```
